---
title: "DATA 605 FUNDAMENTALS OF COMPUTATIONAL MATHEMATICS" 
subtitle: "Assignment 14: Calculus & Taylor Series Expansion"
author: "Kyle Gilde"
date: "12/2/2017"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    toc: true
    toc_depth: 1
---
<style type="text/css">
body{
  font-family: Helvetica;
  font-size: 14pt;
}
</style>
<body>

```{r knitr_options, echo=FALSE}
knitr::opts_chunk$set(
                      error = FALSE
                      ,message = FALSE
                      ,tidy = TRUE
                      #,cache = TRUE
                      )
```


```{r packages, echo=F, collapse=T} 
packages <- c("prettydoc", "Ryacas",'Deriv') 

#see if we need to install any of them
installed_and_loaded <- function(pkg){
  #CODE SOURCE: https://gist.github.com/stevenworthington/3178163
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE, quietly = TRUE, warn.conflicts = FALSE)
}
#excute function and display the loaded packages
data.frame(installed_and_loaded(packages))
```

##Computing the Taylor Series expansion

This week, we'll work out some Taylor Series expansions of popular functions.

$f(x) = 1/(1-x)$

$f(x) = e^x$

$f(x) = ln(1 + x)$

For each function, only consider its valid ranges as indicated in the notes when you are computing the Taylor Series expansion. Please submit your assignment as a R-Markdown document

###$f(x) = 1/(1-x)$

$f(c) = \frac{1}{(1-c)}$

$f'(c) = \frac{1}{(1-c)^2}$

$f''(c) = \frac{2}{(1-c)^3}$

$f'''(c) = \frac{6}{(1-c)^4}$

$f''''(c) = \frac{24}{(1-c)^5}$

\begin{equation}
f(x) =
\frac{1}{(1-c) 0!}  (x-c)^0 + 
\frac{1}{(1-c)^2 1!}  (x-c)^1 + 
\frac{2}{(1-c)^3 2!}  (x-c)^2  + 
\frac{6}{(1-c)^4 3!}  (x-c)^3  + 
\frac{24}{(1-c)^5 4!}  (x-c)^4  + ...
\end{equation}

\begin{equation}
=
\frac{1}{(1-c)} + 
\frac{1}{(1-c)^2} (x-c) + 
\frac{1}{(1-c)^3} (x-c)^2  + 
\frac{1}{(1-c)^4} (x-c)^3  + 
\frac{1}{(1-c)^5} (x-c)^4  +...
\end{equation}

$= \sum_{n=0}^{\infty} \frac{(x-c)^n}{(1-c)^{n+1}}$

**Ratio test:**

\begin{equation}
L = \lim_{n\to\infty}\frac{a_{n+1}}{a_n}
\end{equation}

\begin{equation}
L = lim_{n\to\infty}

\frac{
  \frac{(x-c)^{n+1}}{(1-c)^{n+2}}
}{
  \frac{(x-c)^n}{(1-c)^{n+1}}
}
\end{equation}

\begin{equation}
L = lim_{n\to\infty}
  \frac{(x-c)^{n+1}(1-c)^{n+1}}{(x-c)^n(1-c)^{n+2}}
  = \frac{x-c}{1-c}
  = x
\end{equation}

The series will only converge (L < 1) when the range is between -1 and 1.

###$f(x) = e^x$

$f(c) = e^c$

$f'(c) = e^c$

$f''(c) = e^c$

$f'''(c) = e^c$

$f''''(c) = e^c$

\begin{equation}
f(x) =
\frac{e^c}{0!}  (x-c)^0 + 
\frac{e^c}{1!}   (x-c)^1 + 
\frac{e^c}{2!}   (x-c)^2  + 
\frac{e^c}{3!}   (x-c)^3  + 
\frac{e^c}{4!}   (x-c)^4  + ...
\end{equation}

\begin{equation}
f(x) =
e^c + 
e^c   (x-c) + 
\frac{e^c}{2!}   (x-c)^2  + 
\frac{e^c}{3!}   (x-c)^3  + 
\frac{e^c}{4!}   (x-c)^4  + ...
\end{equation}

$f(x) = \sum_{n=0}^{\infty} \frac{e^c(x-c)^n}{n!}$


**Ratio test:**

\begin{equation}
L = \lim_{n\to\infty}\frac{a_{n+1}}{a_n}
\end{equation}

\begin{equation}
L = lim_{n\to\infty}

\frac{
  \frac{e^c(x-c)^(n+1)}{(n+1)!}
}{
  \frac{e^c(x-c)^n}{n!}
}
\end{equation}

\begin{equation}
L = lim_{n\to\infty}
  \frac{e^c(x-c)^{n+1}n!}{e^c(x-c)^n(n+1)!}
  =  \frac{x-c}{n!}
  = 0
\end{equation}

Since $lim_{n\to\infty} = 0$, the series will converge (L < 1) when the range is $(-\infty, \infty)$.

###$f(x) = ln(1 + x)$

$f(c) = ln(1 + c)$

$f'(c) = 1/(1 + c)$

$f''(c) = -1/(1 + c)^2$

$f'''(c) = 2/(1 + c)^3$

$f''''(c) = -6/(1 + c)^4$

\begin{equation}
f(x) =
\frac{ln(1 + c)}{0!}  (x-c)^0 + 
\frac{1}{(1 + c)1!}   (x-c)^1 + 
\frac{-1}{(1 + c)^2 2!}   (x-c)^2  + 
\frac{2}{(1 + c)^3 3!}   (x-c)^3  + 
\frac{-6}{(1 + c)^4 4!}   (x-c)^4  + ...
\end{equation}

\begin{equation}
f(x) =
ln(1 + c) + 
\frac{x-c}{(1 + c)}    - 
\frac{(x-c)^2}{2(1 + c)^2}     + 
\frac{(x-c)^3}{3(1 + c)^3}    - 
\frac{(x-c)^4}{4(1 + c)^4}     + ...
\end{equation}

$f(x) = ln(1 + c) + \sum_{n=1}^{\infty} (-1)^{n+1}\frac{(x-c)^n}{n(1 + c)^n}$

**Ratio test:**

\begin{equation}
L = \lim_{n\to\infty}\frac{a_{n+1}}{a_n}
\end{equation}

\begin{equation}
L = lim_{n\to\infty}

\frac{
  ln(1 + c) + \frac{(-1)^{n+2}(x-c)^{n+1}}{(n+1)(1 + c)^{n+1}}
}{
  ln(1 + c) + \frac{(-1)^{n+1}(x-c)^n}{n(1 + c)^n}
}
\end{equation}

\begin{equation}
L = 
lim_{n\to\infty} 
1 + \frac{n(1 + c)^n(-1)^{n+2}(x-c)^{n+1}}{(n+1)(1 + c)^{n+1}(-1)^{n+1}(x-c)^n} = 
\end{equation}

\begin{equation}
L = 
lim_{n\to\infty} 
1 - \frac{n(x-c)}{(n+1)(1 + c)} 
= 
lim_{n\to\infty} 
1 - \frac{nx}{n+1} 
= 1 - x
\end{equation}

The series will only converge (L < 1) when the range is between -1 and 1.

###References
http://www.wolframalpha.com/widget/widgetPopup.jsp?p=v&id=f9476968629e1163bd4a3ba839d60925&title=Taylor%20Series&theme=blue&i0=ln(1%20%2B%20x)%20&i1=x&i2=0&i3=10&podSelect=&showAssumptions=1&showWarnings=1
</body>