---
title: "DATA 605 FUNDAMENTALS OF COMPUTATIONAL MATHEMATICS" 
subtitle: "Discussion 13: Calculus Ch. 4 & 7"
author: "Kyle Gilde"
date: "11/19/2017"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    toc: true
    toc_depth: 1
---
<style type="text/css">
body{
  font-family: Helvetica;
  font-size: 14pt;
}
</style>
<body>

```{r knitr_options, echo=FALSE}
knitr::opts_chunk$set(
                      error = FALSE
                      ,message = FALSE
                      ,tidy = TRUE
                      #,cache = TRUE
                      )
```


```{r packages, echo=F, collapse=T} 
packages <- c("prettydoc", "stats", "Deriv")

#see if we need to install any of them
installed_and_loaded <- function(pkg){
  #CODE SOURCE: https://gist.github.com/stevenworthington/3178163
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE, quietly = TRUE, warn.conflicts = FALSE)
}
#excute function and display the loaded packages
data.frame(installed_and_loaded(packages))
```
#Using R, provide the solution for any exercise in either Chapter 4 or Chapter 7 of the calculus textbook.  If you are unsure of your solution, post your concerns.

##Chapter 4.3

3. Find the maximum product of two numbers (not necessarily integers) that have a sum of 100.

```{r}
max_value <- 1000
min_value <- -max_value

fx <- function(x) {x * (100 - x)}
fxprime <-Deriv(fx)
out_root <- uniroot(fxprime, c(min_value, max_value))

fx(out_root$root)

x <- seq(min_value, max_value)
plot(x, fx(x))
abline(out_root$root, out_root$f.root, col = "red")
```


4. Find the minimum sum of two numbers whose product is 500.

$\lim x \rightarrow 0^- f(x) = -\infty$

```{r}
max_value <- 100
min_value <- -max_value

fx <- function(x) {x + 500/x}
fxprime <- Deriv(fx)
(out_root <- uniroot(fx, c(min_value, max_value)))
```


</body>