---
title: "DATA 605 FUNDAMENTALS OF COMPUTATIONAL MATHEMATICS" 
subtitle: "Discussion W6: 3.2 #8 & #10"
author: "Kyle Gilde"
date: "9/27/2017"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    toc: true
    toc_depth: 1
---
<style type="text/css">
body{
  font-family: Helvetica;
  font-size: 14pt;
}
</style>
<body>

```{r, echo = F}
#load_packages <- c("prob")

#see if we need to install any of them
install_load <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE, quietly = TRUE, warn.conflicts = FALSE)
}

#install_load(load_packages)
```

###8. A die is rolled 30 times. What is the probability that a 6 turns up exactly 5 times? 

###In R
```{r, collapse=T}
dbinom(5, 30, 1/6)
```
What is the most probable number of times that a 6 will turn up?
```{r}
30 * 1/6
```

###10. In a ten-question true-false exam, find the probability that a student gets a grade of 70 percent or better by guessing. Answer the same question if the test has 30 questions, and if the test has 50 questions.

```{r}
nquestions <- 10
pct70 <- ceiling(nquestions * .7) - 1
pbinom(pct70, nquestions, .5, lower.tail = F)

nquestions <- 30
pct70 <- ceiling(nquestions * .7) - 1
pbinom(pct70, nquestions, .5, lower.tail = F)

nquestions <- 50
pct70 <- ceiling(nquestions * .7) - 1
pbinom(pct70, nquestions, .5, lower.tail = F)
```


</body>