---
title: "DATA 605 FUNDAMENTALS OF COMPUTATIONAL MATHEMATICS" 
subtitle: "Discussion 14: Calculus Ch. 8.8"
author: "Kyle Gilde"
date: "11/29/2017"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    toc: true
    toc_depth: 1
---
<style type="text/css">
body{
  font-family: Helvetica;
  font-size: 14pt;
}
</style>
<body>

```{r knitr_options, echo=FALSE}
knitr::opts_chunk$set(
                      error = FALSE
                      ,message = FALSE
                      ,tidy = TRUE
                      #,cache = TRUE
                      )
```


```{r packages, echo=F, collapse=T} 
packages <- c("prettydoc")

#see if we need to install any of them
installed_and_loaded <- function(pkg){
  #CODE SOURCE: https://gist.github.com/stevenworthington/3178163
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE, quietly = TRUE, warn.conflicts = FALSE)
}
#excute function and display the loaded packages
data.frame(installed_and_loaded(packages))
```
###Key Idea 32 gives the n th term of the Taylor series of common funcOons. In Exercises 3 - 6, verify the formula given in the Key Idea by finding the first few terms of the Taylor series of the given funcOon and idenifying a pattern.

**4. f(x) = sin x; c = 0**

\begin{equation} \sin x = \sum_{n=0}^{\infty} (-1)^n \frac{x^{2n+1}}{(2n + 1)!} \end{equation}

\begin{equation} f(x) = (-1)^0 \frac{x^{2 * 0 + 1}}{(2*0 + 1)!} + (-1)^1 \frac{x^{2*1+1}}{(2*1 + 1)!} + (-1)^2 \frac{x^{2*2+1}}{(2*2 + 1)!} + (-1)^3 \frac{x^{2*3+1}}{(2*3+1} + ...
\end{equation}

\begin{equation} f(x) = 1 * \frac{x^{1}}{(1)!} + -1 * \frac{x^{3}}{(3)!} + 1 * \frac{x^{5}}{(5)!} + -1 * \frac{x^{7}}{(7)!} + ...
\end{equation}

\begin{equation} f(x) = x - \frac{x^3}{3!} + \frac{x^5}{5!} - \frac{x^7}{7!} + ...
\end{equation}

</body>