---
title: "DATA 605 FUNDAMENTALS OF COMPUTATIONAL MATHEMATICS" 
subtitle: "Assignment 2"
author: "Kyle Gilde"
date: "9/2/2017"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    toc: true
    toc_depth: 1
---
<style type="text/css">
body{
  font-family: Helvetica;
  font-size: 14pt;
}
</style>
<body>
```{r knitr_options, echo=FALSE}
knitr::opts_chunk$set(
                      warning = FALSE
                      ,error = FALSE
                      ,message = FALSE
                      ,tidy = TRUE
                      #,cache = TRUE
                      )
```


```{r packages, echo=F} 
load_packages <- c("prettydoc")

#see if we need to install any of them
install_load <- function(pkg){
  #CODE SOURCE: https://gist.github.com/stevenworthington/3178163
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE, quietly = TRUE, warn.conflicts = FALSE)
}

#library(prettydoc)

#excute function and display the loaded packages
data.frame(t(t(install_load(load_packages))), fix.empty.names = FALSE)
```

##1. Problem set 1

(1) Show that $A^TA \neq AA^T$ in general. (Proof and demonstration.)

**True, in the random sampling of small matrices, the ~86% of the 2 matrix products were not equal.**

```{r}

max_dim <- 3
max_range <- 3
iterations <- 10000

results_df <- data.frame(
  mat = character(), 
  m = integer(),
  n = integer(),
  conjecture_test = logical(),
  matrix_type = character(), 
  stringsAsFactors=F
)

matrix_type_test <- function(mat, tmat, m_dim, n_dim, A_result, B_result){
  if(all(m_dim == 1, n_dim == 1)){
    return("1 by 1 matrix")
  } else if(identical(mat, diag(m_dim))){
    return("Is identify matrix")
  } else if(identical(mat, tmat)){
    return("Is symmetric matrix")
  } else if(m_dim == n_dim & identical(mat %*% tmat, diag(m_dim))){
    return("Is orthogonal matrix")
  } else if(identical(A_result, B_result) & identical(A_result, diag(A_result[1, 1], nrow(A_result)))){
    return("Both products are the same diagonal matrix")
  } else {
    return("N/A")
  }
}



for (i in 1:iterations){

  #select matrix dims
  dim_range <- c(1:max_dim)
  m_dim <- sample(dim_range, 1)
  n_dim <- sample(dim_range, 1)
  samp_size <- m_dim * n_dim

  #select matrix elements
  samp_space <- c(-max_range:max_range)  
  rand_samp <- sample(samp_space, samp_size, replace = T)
  
  #create matrices
  current_matrix <- matrix(rand_samp, m_dim)
  t_current_matrix <- t(current_matrix)
  
  #multiply
  A_result <- t_current_matrix %*% current_matrix
  B_result <- current_matrix %*% t_current_matrix
  
  
  #test & store the results
  match_test <- identical(A_result, B_result)
  matrix_type <- matrix_type_test(current_matrix, t_current_matrix, m_dim, n_dim, A_result, B_result)
  
  new_row <- list(c(paste(rand_samp, collapse = ",")), m_dim, n_dim, match_test, matrix_type)
  
  results_df[i, ] <- new_row
  
}  

prop.table(table(results_df$conjecture_test))

# View(subset(results_df, conjecture_test == TRUE & matrix_type == "N/A"))
# View(subset(results_df, matrix_type == "Is orthogonal matrix"))

#mat <- matrix(c(1,5))

```




(2) For a special type of square matrix A, we get $A^TA = AA^T$. Under what conditions could this be true? (Hint: The Identity matrix I is an example of such a matrix).

**I identified 4 conditions for when the 2 matrix products are equal to each other. **

+ 1 by 1 matrices

+ Symmetric matrices

+ Othogonal matrices

+ When both matrix products produce the same diagonal matrix


```{r}

t(prop.table(table(results_df[, 4:5])))

```



##2. Problem set 2

Matrix factorization is a very important problem. There are supercomputers built just to do matrix  factorizations. Every second you are on an airplane, matrices are being factorized. Radars that track flights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your flight using radars. 

Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer. Please submit your response in an R Markdown document using our class naming convention, E.g. LFulton_Assignment2_PS2.png You don't have to worry about permuting rows of A and you can assume that A is less than 5x5, if you need to hard-code any variables in your code. If you doing the entire assignment in R, then please submit only one markdown document for both the problems.


```{r}
LU_decomp <- function(A){
  size <- nrow(A)
  U <- A
  I <- diag(size)
  myvector <- c()
  
  for (i in 2:size){
    j_max <- i - 1
    for (j in 1:j_max){
      elim_mat <- I
      pivot <- ifelse(U[i - 1, j] != 0, U[i - 1, j], U[i - 2, j])
      elim_mat[i, j] <- - U[i, j] / pivot
      (U <- elim_mat %*% U)
      var_name <- paste0("E", i, j)
      assign(var_name, elim_mat)
      myvector <- c(myvector, var_name)
    }
  }
  
  L <- diag(size)
  for (m in myvector){
    L <- L %*% solve(get(m))
  }
   
  print(U)
  print(L)
  ifelse(identical(L %*% U, A), "The decomposition works", "The decomposition doesn't work")
}


A <- matrix(c(1,2,3,1,1,1,2,0,1),nrow=3)
LU_decomp(A)


A <- matrix(c(1, -2, 3, 4,8,4,-3,5,7), 3)
LU_decomp(A)
```





</body>