---
title: "DATA 605 FUNDAMENTALS OF COMPUTATIONAL MATHEMATICS" 
subtitle: "Homework Week 12: Multiple Linear Regression"
author: "Kyle Gilde"
date: "11/16/2017"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    toc: true
    toc_depth: 2
---
<style type="text/css">
body{
  font-family: Helvetica;
  font-size: 14pt;
}
</style>
<body>

```{r knitr_options, echo=FALSE}
knitr::opts_chunk$set(
                      warning = FALSE
                      ,error = FALSE
                      ,message = FALSE
                      ,tidy = TRUE
                      ,cache = TRUE
                      )
```


```{r packages, echo=F, collapse=T, warning=FALSE} 
packages <- c("prettydoc", "dplyr", "ggplot2", "datasets", "gvlma", "MASS", "lmtest", "car")
installed_and_loaded <- function(pkg){
  #CODE SOURCE: https://gist.github.com/stevenworthington/3178163
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE, quietly = TRUE, warn.conflicts = FALSE)
}
#excute function and display the loaded packages
data.frame(installed_and_loaded(packages))
```

#The Data

The attached who.csv dataset contains real-world data from 2008. The variables included following:

Country: name of the country
LifeExp: average life expectancy for the country in years
InfantSurvival: proportion of those surviving to one year or more
Under5Survival: proportion of those surviving to five years or more
TBFree: proportion of the population without TB.
PropMD: proportion of the population who are MDs
PropRN: proportion of the population who are RNs
PersExp: mean personal expenditures on healthcare in US dollars at average exchange rate
GovtExp: mean government expenditures per capita on healthcare, US dollars at average exchange rate
TotExp: sum of personal and government expenditures.

```{r}
cars_df <- cars
glimpse(cars_df)
```


#Question

**Is the measure of color intensity predictive of the cars's speed content?**

#Model Summary

```{r}
cars_model <- lm(dist ~ speed, cars_df)
summary(cars_model)

intercept <- coef(cars_model)[1]
slope <- coef(cars_model)[2]
```

```{r scatter}
a <- ggplot(cars_model, aes(speed, dist))
a + geom_point() + geom_abline(slope = slope, intercept = intercept, show.legend=TRUE)
```


#Model Interpretation

+ This linear model is expressed as $\widehat{StoppingDistance} = -17.5791 + 3.9324*{speed}$

+ For each additional increase in the miles per hour, the model expects an increase of 3.9 feet in stopping distance.

+ In this model, multiple $R^2$ is 0.6511, which means that the model's least-squares line accounts for approximately $65\%$ of the variation in the stopping distance. 

+ The speed's p-value is near zero and Y-intercept's p-value is ~1%, which means that there is very little chance that they are not relevant to the model.

#Model Diagnostics

Let's assess if this linear model is reliable.

##Linearity: Do the variables have a linear relationship?

The Component+Residual plot shows some deviation from a linear relationship.

```{r}
crPlots(cars_model)
```


##Normality: Are the model's residuals distributed nearly normally?

No, per the histogram and Q-Q plot, the residuals are not normally distributed. There are some outliers on the left-side of the distribution.

```{r hist-res, eval=TRUE}
sresid <- studres(cars_model) 
hist(sresid, freq=FALSE, breaks = 10,
     main="Distribution of Studentized Residuals")
xfit<-seq(min(sresid),max(sresid),length=40) 
yfit<-dnorm(xfit) 
lines(xfit, yfit)
```

```{r qq-res, eval=TRUE}
plot(cars_model, which = 2)
```

##Homoscedasticity: Is there constant variability among the residuals?

+ Based on the scatter plot, the residuals do show some small deviation in variability. 

+ The Non-constant Variance Score Test has a p-value of <.05, which means that we reject the null hypothesis of homoscedasticity.

```{r residuals, eval=TRUE}
plot(cars_model, which = 1)
ncvTest(cars_model) 
```

##Independence: Are the data from a random sample and not from a time series?

The Durbin Watson test's p-value is >.05. Therefore, we fail to reject the null hypothesis of independence (no autocorrelation).

```{r}
durbinWatsonTest(cars_model)
```

#Conclusion

Based upon my diagnostics and the gvlma function, the conditions for linear regression have not been met.

```{r}
gvlma(cars_model)
plot(gvlma(cars_model))
```



#References
[Global Validation of Linear Models Assumptions](https://www.rdocumentation.org/packages/gvlma/versions/1.0.0.2)

http://r-statistics.co/Assumptions-of-Linear-Regression.html

https://www.statmethods.net/stats/regression.html

http://ademos.people.uic.edu/Chapter12.html#121_example_2:_a_thanksgiving_not_to_remember,_part_2

http://rstatistics.net/how-to-test-a-regression-model-for-heteroscedasticity-and-if-present-how-to-correct-it/

http://www.statisticshowto.com/durbin-watson-test-coefficient/

http://www.ianruginski.com/regressionassumptionswithR_tutorial.html

</body>