---
title: "DATA 605 FUNDAMENTALS OF COMPUTATIONAL MATHEMATICS" 
subtitle: "Assignment 2"
author: "Kyle Gilde"
date: "9/2/2017"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    toc: true
    toc_depth: 1
---
<style type="text/css">
body{
  font-family: Helvetica;
  font-size: 14pt;
}
</style>
<body>
```{r knitr_options, echo=FALSE}
knitr::opts_chunk$set(
                      warning = FALSE
                      ,error = FALSE
                      ,message = FALSE
                      ,tidy = TRUE
                      #,cache = TRUE
                      )
```


```{r packages, echo=F} 
load_packages <- c("prettydoc")

#see if we need to install any of them
install_load <- function(pkg){
  #CODE SOURCE: https://gist.github.com/stevenworthington/3178163
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE, quietly = TRUE, warn.conflicts = FALSE)
}

#library(prettydoc)

#excute function and display the loaded packages
data.frame(t(t(install_load(load_packages))), fix.empty.names = FALSE)
```

##1. Problem set 1

(1) Show that $A^TA \neq AA^T$ in general. (Proof and demonstration.)

```{r}
r <- c(-20:20)
mn
d <- sample(r, replace = T)


```


(2) For a special type of square matrix A, we get $A^TA = AA^T$. Under what conditions could this be true? (Hint: The Identity matrix I is an example of such a matrix).


##2. Problem set 2

Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being factorized. Radars that track flights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your flight using radars. 

Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer. Please submit your response in an R Markdown document using our class naming convention, E.g. LFulton_Assignment2_PS2.png You don't have to worry about permuting rows of A and you can assume that A is less than 5x5, if you need to hard-code any variables in your code. If you doing the entire assignment in R, then please submit only one markdown document for both the problems.

```{r}
d <- sample(c(-5 : 5), 9, replace = T)
m <- matrix(d, 3, 3)
m <- diag(1, 3)
upper.tri(m) 
det(m)
mat <- matrix(1:9, 3, 3)
mat[lower.tri(mat)] <- 0


chol(m)

y <- qr(m)
y
```


</body>