---
title: "DATA 605 FUNDAMENTALS OF COMPUTATIONAL MATHEMATICS" 
subtitle: "Assignment 3"
author: "Kyle Gilde"
date: "9/15/2017"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    toc: true
    toc_depth: 1
---
<style type="text/css">
body{
  font-family: Helvetica;
  font-size: 14pt;
}
</style>
<body>
```{r knitr_options, echo=FALSE}
knitr::opts_chunk$set(
                      warning = FALSE
                      ,error = FALSE
                      ,message = FALSE
                      ,tidy = TRUE
                      ,cache = TRUE
                      )
```


```{r packages, echo=F} 
load_packages <- c("prettydoc", "pracma", "ppls")

#see if we need to install any of them
install_load <- function(pkg){
  #CODE SOURCE: https://gist.github.com/stevenworthington/3178163
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE, quietly = TRUE, warn.conflicts = FALSE)
}

#excute function and display the loaded packages
data.frame(t(t(install_load(load_packages))), fix.empty.names = FALSE)
```

##1. Problem set 1

(1) What is the rank of the matrix A?

\begin{equation}A = \begin{bmatrix} 1 & 2 & 3 & 4\\ −1 & 0 & 1 & 3\\ 0 & 1 & −2 & 1\\ 5 & 4 & −2 & −3 \end{bmatrix}\end{equation}


```{r}
d <- c(1, 2, 3, 4, −1, 0, 1, 3, 0, 1, −2, 1, 5, 4, −2, −3)
A <- matrix(d, 4, 4, byrow = T)
rref(A)
Aqr <- qr(A)
Aqr$rank
```

(2) Given an mxn matrix where m > n, what can be the maximum rank? The minimum
rank, assuming that the matrix is non-zero?

**The maximum rank can be no greater than n.**
**The minimum rank is 1.**

(3) What is the rank of matrix B?

\begin{equation}B = \begin{bmatrix} 1 & 2 & 1\\3 & 6 & 3\\ 2 & 4 & 2\\ \end{bmatrix}\end{equation}


```{r}
d <- c(1, 2, 1, 3, 6, 3, 2, 4, 2)
A <- matrix(d, 3, byrow = T)
rref(A)
Aqr <- qr(A)
Aqr$rank
```

##2. Problem set 2

Compute the eigenvalues and eigenvectors of the matrix A. You’ll need to show your work. You’ll need to write out the characteristic polynomial and show your solution.


\begin{equation}A = \begin{bmatrix} 1 & 2 & 3\\0 & 4 & 5\\ 0 & 0 & 6\\ \end{bmatrix}\end{equation}

\begin{equation}x(\lambda I - A) = 0\end{equation}

\begin{equation}det(\lambda I - A) = 0\end{equation}

1. 
\begin{equation}
\begin{array}{|c|} \lambda  \begin{bmatrix} 1 & 0 & 0\\0 & 1 & 0\\ 0 & 0 & 1\\ \end{bmatrix} - \begin{bmatrix} 1 & 2 & 3\\0 & 4 & 5\\ 0 & 0 & 6\\ \end{bmatrix}\end{array} = 0
\end{equation}

2.
\begin{equation}
\begin{array}{|c|} \begin{bmatrix} \lambda & 0 & 0\\0 & \lambda & 0\\ 0 & 0 & \lambda\\ \end{bmatrix} - \begin{bmatrix} 1 & 2 & 3\\0 & 4 & 5\\ 0 & 0 & 6\\ \end{bmatrix}\end{array} = 0
\end{equation}
3.
\begin{equation}
\begin{array}{|c|}
\lambda - 1 & -2 & -3\\
0 & \lambda - 4 & -5\\ 
0 & 0 & \lambda - 6
\end{array} = 0
\end{equation}


4.

\begin{equation}
(\lambda - 1)
\begin{array}{|cc|}
\lambda - 4 & -5\\
0 & \lambda - 6
\end{array}
- (-2)
\begin{array}{|cc|}
0 & -5 \\
0 & \lambda - 6
\end{array}
+ -3
\begin{array}{|cc|}
0 & \lambda - 4 \\
0 & 0
\end{array}
= 0
\end{equation}

5. 
\begin{equation}
(\lambda - 1)((\lambda - 4)(\lambda - 6) - 0) + 2((0)-(0)) - 3((0)-(0)) = 0
\end{equation}

6. **Eigenvalues:**
\begin{equation}
(\lambda - 1)(\lambda - 4)(\lambda - 6) = 0
\end{equation}

\begin{equation}
\lambda = 6, 4, 1
\end{equation}

```{r}
#validate
d <- c(1, 2, 3, 0, 4, 5, 0, 0, 6)
A <- matrix(d, 3, byrow = T)
Aeigen <- eigen(A)
Aeigen$values
```

7. 
\begin{equation}
(\lambda^2 - 5\lambda + 4)(\lambda - 6) = 0
\end{equation}

8.
\begin{equation}
\lambda^3 - 5\lambda^2 + 4\lambda + -6\lambda^2 + 30\lambda -24 = 0
\end{equation}

9. **Characteristic Polynomial:**
\begin{equation}
p_{A}(\lambda) = \lambda^3 - 11\lambda^2 + 34\lambda -24
\end{equation}


```{r validate_cp}
#validate
charpoly(A)
```

10. **To get each eigenvector, let's substitute each $\lambda$ into the matrix from step 3 and find the null spaces**

\begin{equation}
\begin{bmatrix}
\lambda - 1 & -2 & -3\\
0 & \lambda - 4 & -5\\ 
0 & 0 & \lambda - 6
\end{bmatrix}x = 0
\end{equation}

**a.**For $\lambda = 6$

\begin{equation}
\begin{bmatrix}
5 & -2 & -3\\
0 & 2 & -5\\ 
0 & 0 & 0
\end{bmatrix}

\begin{bmatrix}
v_1 \\ v_2 \\ v_3
\end{bmatrix}
=
\begin{bmatrix}
1 & -2/5 & -3/5\\
0 & 1 & -5/2\\ 
0 & 0 & 0
\end{bmatrix}
\begin{bmatrix}
v_1 \\ v_2 \\ v_3
\end{bmatrix}
= 
\begin{bmatrix}
1 & 0 & -8/5\\
0 & 1 & -5/2\\ 
0 & 0 & 0
\end{bmatrix}
\begin{bmatrix}
v_1 \\ v_2 \\ v_3
\end{bmatrix}
= 
\begin{bmatrix}
0 \\ 0 \\ 0
\end{bmatrix}
\end{equation}

\begin{equation}
v_1 = 8/5v_3, v_2 = 5/2v_3
\end{equation}

\begin{equation}
x =
\begin{bmatrix}
v_1 \\ v_2 \\ v_3
\end{bmatrix}
=
v_3\begin{bmatrix}
8/5 \\ 5/2 \\ 1
\end{bmatrix}
=
\begin{bmatrix}
8/5 \\ 5/2 \\ 1
\end{bmatrix}
\end{equation}


\begin{equation}
E_{\lambda=6} = \begin{bmatrix}
8/5 \\ 5/2 \\ 1
\end{bmatrix}
\end{equation}
**Normalized:**

\begin{equation}
||x|| = \sqrt{(8/5)^2 + (5/2)^2 + 1^2} = 3.132092
\end{equation}

\begin{equation}
E_{\lambda=6} = \begin{bmatrix}
0.5108407 \\ 0.7981886 \\ 0.3192754
\end{bmatrix}
\end{equation}
```{r, collapse=T}
#validated
x <- c(8/5, 5/2, 1)
(xlen <- norm(x, type="2"))
x/xlen
all.equal(x/xlen, Aeigen$vectors[, 1])
```


**b.** For $\lambda = 4$

\begin{equation}
\begin{bmatrix}
3 & -2 & -3\\
0 & 0 & -5\\ 
0 & 0 & -2
\end{bmatrix}

\begin{bmatrix}
v_1 \\ v_2 \\ v_3
\end{bmatrix}
=
\begin{bmatrix}
1 & -2/3 & 0\\
0 & 0 & 1\\ 
0 & 0 & 0
\end{bmatrix}
\begin{bmatrix}
v_1 \\ v_2 \\ v_3
\end{bmatrix}
= 
\begin{bmatrix}
0 \\ 0 \\ 0
\end{bmatrix}
\end{equation}

\begin{equation}
v_1 = 2/3v_2, v_3 = 0v_2
\end{equation}

\begin{equation}
x =
\begin{bmatrix}
v_1 \\ v_2 \\ v_3
\end{bmatrix}
=
v_2\begin{bmatrix}
2/3 \\ 1 \\ 0
\end{bmatrix}
=
\begin{bmatrix}
2/3 \\ 1 \\ 0
\end{bmatrix}
\end{equation}


\begin{equation}
E_{\lambda=6} = \begin{bmatrix}
2/3 \\ 1 \\ 0
\end{bmatrix}
\end{equation}
**Normalized:**

\begin{equation}
||x|| = \sqrt{(2/3)^2 + 1^2 + 0^2} = 1.20185
\end{equation}

\begin{equation}
E_{\lambda=6} = \begin{bmatrix}
0.5547002 \\ 0.8320503 \\ 0.0000000
\end{bmatrix}
\end{equation}

```{r, collapse=T}
#validated
d <- c(3, -2, -3, 0, 0, -5, 0, 0, -2)
D <- matrix(d, 3, 3, byrow = T)
rref(D)

x <- c(2/3, 1,  0)
(xlen <- norm(x, type="2") )
x/xlen
all.equal(x/xlen, Aeigen$vectors[, 2])
```

**c.** For $\lambda = 1$

\begin{equation}
\begin{bmatrix}
0 & -2 & -3\\
0 & -3 & -5\\ 
0 & 0 & -5
\end{bmatrix}

\begin{bmatrix}
v_1 \\ v_2 \\ v_3
\end{bmatrix}
=
\begin{bmatrix}
0 & 1 & 0\\
0 & 0 & 1\\ 
0 & 0 & 0
\end{bmatrix}
\begin{bmatrix}
v_1 \\ v_2 \\ v_3
\end{bmatrix}
= 
\begin{bmatrix}
0 \\ 0 \\ 0
\end{bmatrix}
\end{equation}

\begin{equation}
v_2 = 0v_1, v_3 = 0v_1
\end{equation}

\begin{equation}
x =
\begin{bmatrix}
v_1 \\ v_2 \\ v_3
\end{bmatrix}
=
v_1\begin{bmatrix}
1 \\ 0 \\ 0
\end{bmatrix}
=
\begin{bmatrix}
1 \\ 0 \\ 0
\end{bmatrix}
\end{equation}


\begin{equation}
E_{\lambda=6} = \begin{bmatrix}
1 \\ 0 \\ 0
\end{bmatrix}
\end{equation}
**Normalized:**

\begin{equation}
||x|| = \sqrt{1^2 + 0^2 + 0^2} = 1
\end{equation}

\begin{equation}
E_{\lambda=6} = \begin{bmatrix}
1 \\ 0 \\ 0
\end{bmatrix}
\end{equation}

```{r, collapse=T}
#validated
d <- c(0, -2, -3, 0, -3, -5, 0, 0, -5)
D <- matrix(d, 3, 3, byrow = T)
rref(D)

x <- c(1, 0, 0)
(xlen <- norm(x, type="2") )
x/xlen
all.equal(x/xlen, Aeigen$vectors[, 3])
```



Please show your work using an R-markdown document. Please name your assignment submission with your first initial and last name.


</body>