---
title: "DATA 605 FUNDAMENTALS OF COMPUTATIONAL MATHEMATICS" 
subtitle: "Assignment 5"
author: "Kyle Gilde"
date: "10/1/2017"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    toc: true
    toc_depth: 1
---
<style type="text/css">
body{
  font-family: Helvetica;
  font-size: 14pt;
}
</style>
<body>
```{r knitr_options, echo=FALSE}
knitr::opts_chunk$set(
                      warning = FALSE
                      ,error = FALSE
                      ,message = FALSE
                      ,tidy = TRUE
                      #,cache = TRUE
                      )
```


```{r packages, echo=F, collapse=T, warning=FALSE} 
packages <- c("prettydoc", "stats", "rmutil")

#see if we need to install any of them
installed_and_loaded <- function(pkg){
  #CODE SOURCE: https://gist.github.com/stevenworthington/3178163
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE, quietly = TRUE, warn.conflicts = FALSE)
}
#excute function and display the loaded packages
data.frame(installed_and_loaded(packages))
```

###Choose independently two numbers B and C at random from the interval [0, 1] with uniform density. Prove that B and C are proper probability distributions. Note that the point (B,C) is then chosen at random in the unit square.  

Find the probability that

(a) B + C < 1/2.
```{r}
#the integral of the function = 1
fab <- function(a, b){(dunif(a) + dunif(b))}
upper_bound <- .5
lower_bound <- 0
(Fab <- int2(fab, c(lower_bound, lower_bound), c(upper_bound, upper_bound)))
Fab == 1

#non-negativity
x <- runif(1000)
df <- expand.grid(x, x)
min(fab(df$Var1, df$Var2)) > 0


#probability
set.seed(.6)
B <- runif(100000)
C <- runif(100000)

sum(B + C < .5) / length(B)
```



(b) BC < 1/2.
```{r}

#the integral of the function = 1
fab <- function(a, b){(dunif(a) * dunif(b))}
upper_bound <- 1
lower_bound <- 0
(Fab <- int2(fab, c(lower_bound, lower_bound), c(upper_bound, upper_bound)))
Fab == 1

#non-negativity
x <- runif(1000)
df <- expand.grid(x, x)
min(fab(df$Var1, df$Var2)) > 0

#probability
set.seed(.6)
B <- runif(100000)
C <- runif(100000)

sum(B * C < .5) / length(B)
```


(c) |B − C| < 1/2.
```{r}
#the integral of the function = 1
fab <- function(a, b){abs(dunif(a) - dunif(b))}
upper_bound <- 1
lower_bound <- 0
(Fab <- int2(fab, c(lower_bound, lower_bound), c(upper_bound, upper_bound)))
Fab == 1

#non-negativity
x <- runif(1000)
df <- expand.grid(x, x)
min(fab(df$Var1, df$Var2)) > 0

#probability
set.seed(.6)
B <- runif(100000)
C <- runif(100000)

sum(abs(B - C) < .5) / length(B)
```


(d) max{B,C} < 1/2.
```{r}
#the integral of the function = 1
fab <- function(a, b){max(dunif(a), dunif(b))}
upper_bound <- 1
lower_bound <- 0
(Fab <- int2(fab, c(lower_bound, lower_bound), c(upper_bound, upper_bound)))
Fab == 1

#non-negativity
x <- runif(1000)
df <- expand.grid(x, x)
min(fab(df$Var1, df$Var2)) > 0

#probability
set.seed(.6)
B <- runif(100000)
C <- runif(100000)

sum(pmax(B, C) < .5) / length(B)
```


(e) min{B,C} < 1/2.
```{r}
#the integral of the function = 1
fab <- function(a, b){min(dunif(a), dunif(b))}
upper_bound <- 1
lower_bound <- 0
(Fab <- int2(fab, c(lower_bound, lower_bound), c(upper_bound, upper_bound)))
Fab == 1

#non-negativity
x <- runif(1000)
df <- expand.grid(x, x)
min(fab(df$Var1, df$Var2)) > 0


#probability
set.seed(.6)
B <- runif(100000)
C <- runif(100000)

sum(pmin(B, C) < .5) / length(B)
```


</body>