---
title: "DATA 605 FUNDAMENTALS OF COMPUTATIONAL MATHEMATICS" 
subtitle: "Assignment 13: Calculus"
author: "Kyle Gilde"
date: "11/26/2017"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    toc: true
    toc_depth: 1
---
<style type="text/css">
body{
  font-family: Helvetica;
  font-size: 14pt;
}
</style>
<body>

```{r knitr_options, echo=FALSE}
knitr::opts_chunk$set(
                      error = FALSE
                      ,message = FALSE
                      ,tidy = TRUE
                      #,cache = TRUE
                      )
```


```{r packages, echo=F, collapse=T} 
packages <- c("prettydoc", "Deriv","rootSolve","rSymPy") 

#see if we need to install any of them
installed_and_loaded <- function(pkg){
  #CODE SOURCE: https://gist.github.com/stevenworthington/3178163
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE, quietly = TRUE, warn.conflicts = FALSE)
}
#excute function and display the loaded packages
data.frame(installed_and_loaded(packages))
```

###1. Use integration by substitution to solve the integral below.

$\int 4e^{-7x} dx$

```{r}
x <- Var("x")
sympy("integrate(4 * exp(-7 * x))")
```


###2. Biologists are treating a pond contaminated with bacteria. The level of contamination is changing at a rate of $dN/dt = -3150/t^4 - 220$ bacteria per cubic centimeter per day, where t is the number of days since treatment began. Find a function N( t ) to estimate the level of contamination if the level after 1 day was 6530 bacteria per cubic centimeter.

```{r}
ft <- function(t) -3150 * t^-4 - 220 
#anti-derivative

#F(t) = 1050 * t^-3 - 220 * t + C

#solve for C: 

#F(1) = 6530 = 1050 * t^-3 - 220 * t + C
t <- 1
(C <- 6530 - 1050 * t^-3 + 220 * t)
#Integral
(Ft <- function(t) 1050 * t^-3 - 220 * t + 5700)
```


###3. Find the total area of the red rectangles in the figure below, where the equation of the line is $f(x) = 2x-9$ 



```{r}
fx <- function(x) 2*x-9
#f(0) = 2x-9

#x = 4.5
(lower <- 4.5)
(upper <- lower + 4)
stats::integrate(Vectorize(fx), lower, upper)
```

###4. Find the area of the region bounded by the graphs of the given equations.

$y = x^2 -2x -2$

$y = x + 2$

```{r}
s <- seq(-2, 5, by=.01)
fx <- function(x) x^2 -2*x -2
fx2 <- function(x) x + 2

fx_values <- fx(s)
fx2_values <- fx2(s)

plot(s, fx_values,type="l", col="red")
lines(s,fx2_values,col="green", type = "l")
abline(0,0)

#intersections
is_equal <- fx2_values == fx_values
mydf <- data.frame(cbind(s, is_equal))
intersections <- subset(mydf, is_equal==1)

#critical points
roots <- uniroot.all(fx, c(-2, 5))
lower1 <- roots[1]
upper1 <- roots[2]
lower2 <- intersections$s[1]
upper2 <- intersections$s[2]

#answer
(integrate(Vectorize(fx2),lower2, upper2)$value - integrate(Vectorize(fx),lower2, upper2)$value)

###Validate answer###
Fx_neg <- abs(integrate(Vectorize(fx),lower1, upper1)$value)
Fx_pos1 <- integrate(Vectorize(fx),lower2, lower1)$value
Fx_pos2 <- integrate(Vectorize(fx),upper1, upper2)$value
Fx2 <- integrate(Vectorize(fx2),lower2, upper2)$value
Fx_neg + Fx2 - Fx_pos2 - Fx_pos1
```

###5. A beauty supply store expects to sell 110 flat irons during the next year. It costs \$3.75 to store one flat iron for one year. There is a fixed cost of $8.25 for each order. Find the lot size and the number of orders per year that will minimize inventory costs.

```{r 21}
#lotsize  = 110 / orders_per_year

#cost = lotsize*3.75 + orders_per_year * 8.25

total_cost <- function(orders_per_year) 110 / orders_per_year * 3.75 + 8.25 * orders_per_year
(total_cost.prime <- Deriv(total_cost))
 
(root <- uniroot.all(total_cost.prime, c(1, 110)))

#calculate in whole number terms
orders_per_year <- floor(root)
lot_size <- ceiling(110/orders_per_year)
total_units <- orders_per_year * lot_size
min_cost <- lot_size  * 3.75 + 8.25 * orders_per_year

paste(orders_per_year, "orders of", lot_size, "units each,", total_units,  "total units, for a minimum cost of", min_cost)
```


###6. Use integration by parts to solve the integral below.

$\int ln( 9x ) * x^6 dx$

```{r}
x <- Var("x")
sympy("integrate(log( 9 * x ) * x**6)")
```


###7. Determine whether f ( x ) is a probability density function on the interval [1, e^6] . If not, determine the value of the definite integral.
$f ( x ) = 1/6x$
```{r}
fx <- function(x) 1/(6*x)
lower <- 1
upper <- exp(6)

all.equal(1, integrate(Vectorize(fx),lower, upper)$value,
          tolerance = 10^-8)
```


</body>