---
title: "DATA 605 FUNDAMENTALS OF COMPUTATIONAL MATHEMATICS" 
subtitle: "Assignment 15: Calculus Week 3"
author: "Kyle Gilde"
date: "12/6/2017"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    toc: true
    toc_depth: 1
---
<style type="text/css">
body{
  font-family: Helvetica;
  font-size: 14pt;
}
</style>
<body>

```{r knitr_options, echo=FALSE}
knitr::opts_chunk$set(
                      error = FALSE
                      ,message = FALSE
                      ,tidy = TRUE
                      #,cache = TRUE
                      )
```


```{r packages, echo=F, collapse=T} 
packages <- c("prettydoc",'Deriv', 'rootSolve') 

#see if we need to install any of them
installed_and_loaded <- function(pkg){
  #CODE SOURCE: https://gist.github.com/stevenworthington/3178163
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE, quietly = TRUE, warn.conflicts = FALSE)
}
#excute function and display the loaded packages
data.frame(installed_and_loaded(packages))
```

1. Find the equation of the regression line for the given points. Round any final values to the nearest hundredth, if necessary.
( 5.6, 8.8 ), ( 6.3, 12.4 ), ( 7, 14.8 ), ( 7.7, 18.2 ), ( 8.4, 20.8 )

```{r}
given_points <- c(5.6, 8.8 ,  6.3, 12.4 ,  7, 14.8 ,  7.7, 18.2 ,  8.4, 20.8)
D <- matrix(d, ncol = 2, byrow = TRUE)

df <- data.frame(D)
names(df) <- c('x', 'y')

the_model <- lm(y ~ x, df)

b <- the_model$coefficients[1]
m <- the_model$coefficients[2]

paste0('y = ', round(m, 2), 'x + ', b)
```


2. Find all local maxima, local minima, and saddle points for the function given below. Write your answer(s) in the form ( x, y, z ). Separate multiple points with a comma. 

$f ( x, y ) = 24x - 6xy^2 - 8y^3$

```{r}
interval <- 10000

fxy <- function(x, y) {24*x - 6*x*y^2 - 8*y^3}
(fxy_prime <- Deriv(fxy))
(fxy_dblprime <- Deriv(fxy_prime))

f_x <- function(y) 24 - 6 * y^2
f_y  <- function(x, y) -(y * (12 * x + 24 * y))

(f_x_critical_points <- uniroot.all(f_x, c(-interval, interval)))


```



3. A grocery store sells two brands of a product, the "house" brand and a "name" brand. The manager estimates that if she sells the "house" brand for x dollars and the "name" brand for y dollars, she will be able to sell $81 - 21x + 17y$ units of the "house" brand and $40 + 11x - 23y$ units of the "name" brand.

Step 1. Find the revenue function R ( x, y ).

$R (x, y) = x(81 - 21x + 17y) + y(40 + 11x - 23y)$

$R (x, y) = -21x^2 + 81x + 28xy + 40y - 23y^2$

Step 2. What is the revenue if she sells the "house" brand for $2.30 and the "name" brand for $4.10?

```{r}
x <- 2.30
y <- 4.10
fxy <- function(x, y) -21*x^2 + 81*x + 28*x*y + 40*y - 23*y^2

fxy(x, y)
```



4. A company has a plant in Los Angeles and a plant in Denver. The firm is committed to produce a total of 96 units of a product each week. The total weekly cost is given by $C(x, y) = 1/6x^2 + 1/6y^2 + 7x + 25y + 700$, where x is the number of units produced in Los Angeles and y is the number of units produced in Denver. How many units should be produced in each plant to minimize the total weekly cost?

$96 = x + y \rightarrow y = 96 - x$

$1/6x^2 + 1/6y^2 + 7x + 25y + 700 \rightarrow 1/6x^2 + 1/6(96 - x)^2 + 7x + 25(96 - x) + 700$


```{r}
Cx <- function(x) 1/6*x^2 + 1/6*(96 - x)^2 + 7*x + 25*(96 - x) + 700
Cx_prime <- Deriv(Cx)

s <- seq(-100, 100, by = 1)
plot(s, Cx_prime(s), type = "l", col = "red")
abline(0, 0)

(x <- uniroot(Cx_prime, c(-100, 100))$root)
(y <- 96 - x)
```


5. Evaluate the double integral on the given region. Write your answer in exact form without decimals

$\iint(e^{8x + 3y dA})$

$R: 2 \leq x \leq 4 $
and
$2\leq  y \leq 4$


</body>