---
title: "DATA 605 FUNDAMENTALS OF COMPUTATIONAL MATHEMATICS" 
subtitle: "Assignment 7: Random Variables"
author: "Kyle Gilde"
date: "10/14/2017"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    toc: true
    toc_depth: 1
---
<style type="text/css">
body{
  font-family: Helvetica;
  font-size: 14pt;
}
</style>
<body>
```{r knitr_options, echo=FALSE}
knitr::opts_chunk$set(
                      warning = FALSE
                      ,error = FALSE
                      ,message = FALSE
                      ,tidy = TRUE
                      ,cache = TRUE
                      )
```


```{r packages, echo=F, collapse=T, warning=FALSE} 
packages <- c("prettydoc")

#see if we need to install any of them
installed_and_loaded <- function(pkg){
  #CODE SOURCE: https://gist.github.com/stevenworthington/3178163
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE, quietly = TRUE, warn.conflicts = FALSE)
}
#excute function and display the loaded packages
data.frame(installed_and_loaded(packages))
```

###1. Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi's. Find the distribution of Y.

order statistic: distribution of the minimum
https://docs.google.com/document/d/1PqAoVuau_3if7ZJzd7dv8cNPT6n5PNjruk8Hulwjb8c/edit#heading=h.u8c5sr7jr4cj

###2. Your organization owns a copier (future lawyers, etc.) or MRI (future doctors).This machine has a manufacturer's expected lifetime of 10 years. This means that we expect one failure every ten years. (Include the probability statements and R Code for each part.).



+ a. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a geometric. (Hint: the probability is
equivalent to not failing during the first 8 years..)

 \begin{equation}
P(X > 8) = \sum_{x>8}.1(1 - .1)^{x - 1} = 0.4493288
\end{equation}
 \begin{equation}
E(X) = 1/.1 = 10
\end{equation}
 \begin{equation}
\sigma = \sqrt{\frac{1 - .1}{.1^2}} = 9.486833
\end{equation}

```{r, collapse=T}
p <- .1
first_success <- 8
initial_failures <- first_success - 1

pgeom(initial_failures, p, lower.tail = F)
CDF <- 1 - (1 - p)^first_success
1 - CDF

(EX <- 1/p)

(sigma <- sqrt((1 - p)/p^2))



#Convert to minutes for N, pi, t
(N <- 10*365.25*24*60)

(pi <- 1/N)

(t <- 8*365.25*24*60)
#Geometric.
#P(X > t | pi) = 1 - P(X<=t | pi) #0.4493288
1 - pgeom(t, pi)

```

+ b. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as an exponential.

 \begin{equation}
P(X > 8) = \int_{8}^{\infty} \lambda e^{-\lambda x} = 0.449329
\end{equation}
 \begin{equation}
E(X) = 10 
\end{equation}
 \begin{equation}
\sigma = \sqrt{1 / .1^2} = 10
\end{equation}


```{r, collapse=T}
x <- 8
lambda <- .1

pexp(x, lambda, lower.tail = F)
CDF <- 1 - exp(-lambda * x)
1 - CDF

(EX <- 1 / lambda)

(sigma <- sqrt(1 / lambda^2))

#Exponential.  P(X >= t | lambda = pi) = 1 - P(X <= t | pi)
1 - pexp(t, pi)  #0.449329
```

+ c. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8
years)

 \begin{equation}
P(X > 8) = \left( \begin{array}{c} 8 \\ 0 \end{array} \right).1^x(1 - .1)^{8-0} = 0.4493289
\end{equation}
 \begin{equation}
E(X) = 8 * .1 = .8
\end{equation}
 \begin{equation}
\sigma = \sqrt{8 * .1 (1 - .1)} = \sqrt{0.72} = 0.8485281
\end{equation}

```{r, collapse=T}
p <- .1
n <- 8
x <- 0

dbinom(x, n, p)
choose(p, x) * p^x * (1 - p)^(n - x)

(EX <- n * p)

(sigma <- sqrt(n * p * (1 - p)))

#Binomial.
#P(X = 0 | N = t, pi = pi) # 0.4493289
dbinom(0, t, pi)  
```


+ d. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a Poisson. 

 \begin{equation}
P(X > 8) = \sum_{x>8}\frac{10^xe^{-10}}{x!} = 0.449329
\end{equation}
 \begin{equation}
E(X) = .8 
\end{equation}
 \begin{equation}
\sigma = \sqrt{.8} = 0.894427
\end{equation}

```{r, collapse=T}
x <- 0
T <- 8
p <- .1
lambda <- T * p
dpois(x, lambda)
lambda^x * exp(-lambda)/factorial(x)

(EX <- lambda)

(sigma <- sqrt(lambda))

#Poisson.  P(X = 0 | lambda = pi, t)
dpois(0, pi * t)  #0.449329
```

</body>