---
title: "DATA 605 FUNDAMENTALS OF COMPUTATIONAL MATHEMATICS" 
subtitle: "Assignment 6: Conditional, Combinatoric & Bayesian Probability "
author: "Kyle Gilde"
date: "10/7/2017"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    toc: true
    toc_depth: 1
---
<style type="text/css">
body{
  font-family: Helvetica;
  font-size: 14pt;
}
</style>
<body>
```{r knitr_options, echo=FALSE}
knitr::opts_chunk$set(
                      warning = FALSE
                      ,error = FALSE
                      ,message = FALSE
                      ,tidy = TRUE
                      ,cache = TRUE
                      )
```


```{r packages, echo=F, collapse=T, warning=FALSE} 
packages <- c("prettydoc", "prob")

#see if we need to install any of them
installed_and_loaded <- function(pkg){
  #CODE SOURCE: https://gist.github.com/stevenworthington/3178163
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE, quietly = TRUE, warn.conflicts = FALSE)
}
#excute function and display the loaded packages
data.frame(installed_and_loaded(packages))
```

###1. Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi’s. Find the distribution of Y .



###2. Your organization owns a copier (future lawyers, etc.) or MRI (future doctors).
This machine has a manufacturer’s expected lifetime of 10 years. This means
that we expect one failure every ten years. (Include the probability statements
and R Code for each part.).



a. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a geometric. (Hint: the probability is
equivalent to not failing during the first 8 years..)

```{r}
pgeom(7, .1, lower.tail = F)
```



b. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as an exponential.


c. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8
years)

d. What is the probability that the machine will fail after 8 years?. Provide also the
expected value and standard deviation. Model as a Poisson. 

</body>